Investigating the Applications and Limits of Single Asymmetric Particle Light Scattering using Limited Detection Schemes

### 2022-11-14 Manuscript.tex
Add files via upload

\usepackage{amssymb, amsthm}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{wrapfig}
\usepackage{geometry}
\usepackage{fancyref}
\usepackage{lineno}
\usepackage{ragged2e}
\usepackage{xcolor}

@@

\begin{abstract}
\justifying
While optical trapping is a well understood method for force transduction and detection, \textcolor{blue}{ further} characterisation of trapped entities \textcolor{blue}{with complex structure and dynamics} \textcolor{blue}{ would provide valuable information for example on aggregation processes, microscale biological processes and so on. Such characterisation using light scattering from the trapped object} poses a two-fold challenge - one experimental, concerning the arrangement of scattered light detectors, and the other theoretical, involving solution of the inverse light scattering problem. \textcolor{blue} {Experimentally,} combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup;  \textcolor{blue} {while the theoretical inverse scattering problem is a non-trivial task given the potentially complex structure of the entity being characterised. To address these issues} we propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme, \textcolor{blue}{combined with}  a Bayesian inference-based analysis method demonstrating how data simulated to mimic light scattering signals in such a scenario may be used to solve the inverse light scattering problem. \textcolor{blue}{As an example problem} we discuss the application of our method to infer the instantaneous orientation of a simulated trapped asymmetric microsphere dimer from the scattered light signal. We argue that this method can be extended for determining in principle any characteristics of the trapped microstructure that influence the light scattering pattern.
\end{abstract}

@@

@@

Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few piconewtons. The ability to characterise the trapped entity has provided useful insights in several fields of research; such as metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}. Chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering \cite{gupta2014raman}, while dynamical characterisation has been demonstrated using a Quadrant Photo Detector,  by following the centre-of-mass Brownian motion of the trapped entity  \cite{friedrich2012tuning} and measuring rotation  of the centre-of-mass \cite{yifat2021facile}.

The possibility of integrating light scattering with optical trapping was demonstrated by Saffran and co-workers in \cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and its Brownian motion was studied. While this allowed for studying the dynamics, structural information about the trapped bead was precluded as the measurement was obtained only at a single angle. Furthermore, this method cannot be extended for trapped entities that would scatter light non isotopically if their motion is large compared to the dimensions of the trap focus. In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure~\ref{fig:setup}, combined with a novel Bayesian Inference based analysis technique to enable interpretation of the resulting multi-angle data as well as optimisation of the setup to provide maximal information from the signal.

To demonstrate the analysis, we study a simple paradigmatic example of a trapped entity with more complex dynamics and scattering than a single sphere, i.e. a trapped asymmetric dimer; and we explore how to optimise the experimental measurement to obtain the best estimate of the dimer's instantaneous orientation, as a further paradigmatic example of interpretation of the scattering data to obtain dynamical information on the trapped entity. There is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved. Here, we simulate the Brownian motion of a dimer and calculate the resulting scattered light, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation form the light scattering signals given a choice of scattering measurement angles.

\textcolor{blue} {not very happy with the above but I am trying to set out clearly in advance the narrative of what we are doing, starting with the general problem, moving to the specific example we are studying, and introducing how we then go about it. we need to emphasise how oreitnation of a dimer is our choice of example problem, where a dimer is a paradigm of a 'complex object' and its orientation is a paradigm of 'data' that in some way describes it}

@@

\section{Methodology}
\label{sec:Method}
\subsection{Brownian Simulation}
\label{sec:2.1}
We use the Brownian OT package developed by Jerome Fung \textit{et al} \cite{Vigilante2020Brownian_OT} to simulate the motion of a asymmetric dimer within an optical trap. Brownian OT is unique in that it combines MSTM \cite{Mishchenko1996MSTM} and 'Optical Tweezer Toolbox' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary shaped sphere clusters. We simulated the motion of silica dimer trapped in a 5 mW Gaussian beam; the dimer was comprised of two spheres with radii $1\mu m, \ \& \ 0.5 \ \mu m$ respectively. We ran our simulation over the span of 10 seconds, with Brownian OT calculating the dimer's orientation and position every $1 \ ms$.

With any trapped particle there are two main sources of motion, Brownian and optical forces. The dimer's Brownian motion ($\Delta q^{\textbf{B}}$) is simulated by generating a set of normally-distributed numbers whose covariance is given by the dimer's diffusion tensor $\textbf{D}$. The diffusion tensor is computed based on Farsund's work \cite{Farsund1996}. For computing the optical force Brownian OT uses the T-Matrix method to calculate the scattering coefficients of the outgoing scattered wave. The T-matirx describes the scattering process by relating the scattering wave to the incident wave in an elegant equation.

@@

MSTM computes the T-matrix of the dimer, and then passes the T-matrix to \textit{ott} to compute the optical forces. The benefit of this approach is that MSTM can accurately compute the T-matrix for any arbitrary collection of spheres once and then \textit{ott} can calculate the optical forces and toques. The total displacement is the sum of the optical and Brownian contributions.

@@

Where \textbf{F} is the a 6 component vector that describes the optical force and torque components. Brownian OT returns the dimers motion as a \textit{csv} file; with the $[x,y,z]$ position being relative to the trap focus, and the orientation being returned as a quaternion.

\subsection{Theory}
\label{sec:2.2}

If we consider our simulated dimer from before; as it moves within the trap its instantaneous orientation can be represented as a three component unit vector, denoted as $\hat{s}$, one that points from the centre of the larger sphere, to the centre of the smaller sphere. If we shine a plane wave laser, perpendicular to the trapping laser, at the dimer a scattering pattern will be produced that is dependent upon the dimer's instantaneous orientation, denoted as $I(\hat{s},\ \theta)$. This scattering pattern was computed using MSTM for each time step of our Brownian simulation. From the scattering pattern, three angles ($\theta_1,\ \theta_2,\ \& \theta_3$) were chosen and the intensity at each angle k was recorded as $I(\hat{s},\ \theta_k)$.

Instead of trying to get an exact estimation of $\hat{s}$ we discretise the orientation space into 30 possible reference orientations $\hat{n}_i$ that divide up a unit sphere evenly \cite{Rey2006}. Using MSTM we compute the scattering patterns produced by our dimer in that orientation, denoted as $I(\hat{n}_i, \ \theta)$. We use the same choice of angles to record $I(\hat{n}_i, \ \theta_k)$. These raw intensities can vary by several orders of magnitude across the entire range of possible angles, to correct this each intensity is scaled according to \eqref{eq:scale}:

\begin{align}
	\label{eq:scale}
	y_k(\hat{n}_i,\ \theta_k) = 
	 \frac{\left[I(\hat{n}_i, \ \theta_k)- \langle I(\hat{n}_i, \ \theta_k) \rangle \right]} 
	{\langle I^2(\hat{n}_i,\ \theta_k) \rangle -\langle I(\hat{n}_i, \ \theta_k)\rangle^2}
\end{align}

Where the denominator is simply the standard deviation of $I(\hat{n}_i, \theta_k)$; our dimer's instantaneous intensity is also scaled using \eqref{eq:scale} to give $y(\hat{s},\ \theta_k)$. See Table's A.1 and A.2 for numerical values of our reference orientation, raw intensities, and scaled signals. In practice $y(\hat{s}\, \theta_k)$ will have an inherent amount of uncertainty associated with it. We compare our instantaneous signal to our reference signal by assuming that our values of $y(\hat{n}_i,\ \theta_k)$ are normally disturbed at each angle k. Combining all three detection angles provides us with the conditional probability that our dimer in orientation $\hat{n}_i$ will produce a signal $y(\hat{s})$:

\begin{align}
	\label{eq:conditional}
	p_i(y(\hat{s})\parallel\hat{n}_i) = \Pi^3_{k=1}
	(2\pi\sigma&_k^2)^{-1/2}  \nonumber \\
	&e^{-(y_{k}(\hat{n}_i)-y_{k}(\hat{s}))^2/2\sigma_k^2}
\end{align}

Where $\sigma_k$ is the average error in our signal collection; we left this as a tuneable variable to measure how the influence of experimental error would effect our predictive model. The signal error was simulated in our model by stochastically adjusting our values of $y_k(\hat{s})$.

@@

Where $\epsilon$ is the percent error expected from our detection fibres. Equation~\eqref{eq:conditional} is assigning a probability to each reference orientation based on how similar it's signal is to the signal our dimer produced. If however, our reference signals are similar to one another it can be difficult to parse the correct reference orientation. Bayes Theorem was utilised to update our conditional probability by accounting for the priori distribution of reference orientations and instantaneous signals.

@@

Where $p_i(\hat{n})$ and $p(y)$ are our estimations of the priori distributions of particle orientations and instantaneous signals respectively. The former can be estimated by some Boltzmann distribution based on the physical spacing between reference orientations, covered in Section~\ref{sec:2.3}. The latter priori $p(y)$ is the probability of our dimer producing the signal $y(\hat{s})$ and is given as.

@@

The implementation of \eqref{eq:bayes} improves our estimate by now accounting for how our reference orientations are distributed in physical space. In order to evaluate the accuracy of our model we tested it using our Brownian simulation of the dimer, see Section~\ref{sec:2.1}.

@@

\section{Discussion}
\label{sec:Discussion}
\subsection{Asymmetric dimer dynamics}
\label{sec:3.1}
The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\mu m, a_2=0.5\mu m$) over the first 10 seconds of entering the optical trap. The initial orientation was assumed as strictly vertical (in line with the beam propagation direction). The dimer's position orientation was recorded every $10 \mu s$ for using as a test dataset for our model.

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[width =\textwidth]{./Images/traj.png}
		\subcaption{}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\includegraphics[width=\textwidth]{./Images/pos.png}
		\subcaption{}
	\end{subfigure}
	\caption{Simulation results of: a) the dimers orientation vector with time, b) the dimer's [x,y,z] position with time.}
	\label{fig:motion}
\end{figure}

@@

\subsection{Impact of signal error on model predictions}
\label{sec:3.2}
Signal error is an important parameter to control when it comes to understanding how well our model can predict the dimers motion. To this end we ran our model while varying the value of $\epsilon$. \figurename{ 3} shows the model's performance with a selection of angles $15^{\circ},\ 55^{\circ}, \ \& \ 90^{\circ}$ and $\beta$ set at 1. The dotted line indicates the minimum radian distance ($0.137$ radians) between two neighbouring reference orientations, if we are under this line then we know our model is just off from selecting the correct orientation

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\textwidth]{./Images/epsilon_5.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\textwidth]{./Images/epsilon_15.png}
	\end{subfigure}
	\begin{subfigure}{0.32\textwidth}
		\includegraphics[width=\textwidth]{./Images/epsilon_25.png}
	\end{subfigure}
	\caption{Model prediction for signal error of a) $5\%$, $15\%$, and $25\%$}
	\label{fig:epsilon}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{./Images/beta_5.png}
	\end{subfigure}
	\begin{subfigure}{0.3\textwidth}
		\includegraphics[width=\textwidth]{./Images/beta_6.png}
	\end{subfigure}
	\caption{Model prediction for signal error of $25\%$ when a) $\beta = 5$ and b) $\beta=6$}
	\label{fig:beta}
\end{figure}

As shown by \ref{fig:beta} eventually $\beta$ begins to over correct our estimate to the extent that we are a full radian away from the correct result. The exact value of $\beta$ that will optimise our performance varies for each choice of angle, therefore we can use it as variable for finely tuning our results to suit our choice of angles and our signal error. In the future we want to also investigate the possibility of extending our priori estimation so we account for not just the change in orientation after one time step but after 2 or 3 time steps to investigate if we can investigate the auto correlation of our dimer's motion.

\subsection{Ultra-nest Results}
\label{sec:3.3}
We wanted to see if our choice of angles plays a significant role in our model's efficacy. We employed ultra-nest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$. We kept both $\beta$ and $\sigma$ constant during this so we could as these are parameters that can be changed more freely.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.4\textwidth]{./Images/corneranglesfreed-1.png}
	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement}
	\label{fig:corner}
\end{figure}

Figure~\ref{fig:corner} is the resulting corner plot from our ultra-nest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer. Overall from the corner plot provided it's clear that there is no direct correlation between our model's performance and the choice of detection angle. While the choice of angles does not provide a direct improvement in our estimation it does determine how finely we can estimate the orientation, for choices of angles that are close together the minimum spacing between our signals is so small that we can't achieve any greater accuracy than dividing the orientation space into 30 reference orientations. We hope to develop our model further in the future by being able to determine the maximum accuracy given that we know the signal error of our measurements.

@@

\section{Conclusion}
\label{4}
We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern. In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity. The MSTM package is flexible in calculating the light scattering for a variety of micro particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy. This has potential to be applied to particles that are changing while in the optical trap. We hope to apply this work in the future to  characterise processes such as nucleation within an optical trap.\\

\noindent \textbf{Funding.} \\

>>>>

\usepackage{amssymb, amsthm}
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage[T1]{fontenc}
\usepackage{babel}
\usepackage{wrapfig}
\usepackage{geometry}
\usepackage{fancyref}
\usepackage{lineno}
\usepackage{color}
\usepackage{nomecl}

@@

\begin{abstract}
\justifying
While optical trapping is a well understood method for force transduction and detection,characterisation of trapped entities poses a two-fold challenge - one experimental concerning the arrangement of light detectors and the other, theoretical involving solving of the inverse light scattering problem. Combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup. We propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme and demonstrate how a Bayesian inference based analysis of the data simulated to mimic light scattering detection signals in such scenarios maybe used for solving the inverse light scattering problem and help characterise trapped entities. To this end, we discuss the application of our method to infer the instantaneous orientations of an asymmetric microsphere dimer being trapped. We argue that this method can be extended for determining any characteristics of the trapped microstructure that influence the light scattering pattern.
\end{abstract}

@@

\nomenclature{$\hat{s}$}{Dimer's instantaneous orientation}
\nomenclature{$\hat{n}_i$}{Reference Orientation i}

@@

Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics  \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few piconewtons. \textit{Further characterisation of the trapped entity would provide useful information, as explored in for example} metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}. Chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering \cite{gupta2014raman}, while dynamical characterisation has been demonstrated using a Quadrant Photo Detector,  by following the centre-of-mass Brownian motion of the trapped entity  \cite{friedrich2012tuning} and measuring rotation  of the centre-of-mass \cite{yifat2021facile}.

The possibility of integrating light scattering with optical trapping was demonstrated by Saffran and co-workers in \cite{Bar-Ziv_1998} where a single-mode optical fibre was aligned to detect the scattered light from a trapped bead and its Brownian motion was studied. While this allowed for studying the dynamics, structural information about the trapped bead was precluded as the measurement was obtained only at a single angle. In this work, we propose a scheme that expands on the technique in \cite{Bar-Ziv_1998} to detect scattered light simultaneously at 3 angles as shown in Figure~\ref{fig:setup}, combined with a novel Bayesian Inference based analysis technique to enable interpretation of the resulting multi-angle data as well as optimisation of the setup to provide maximal information from the signal.

To demonstrate the analysis, we study a simple paradigmatic example of a trapped entity more complex than a single sphere, i.e. a trapped asymmetric dimer; and we explore how to optimise the experimental measurement to obtain the best estimate of the dimer's instantaneous orientation, as a further paradigmatic example of interpretation of the scattering data to obtain dynamical information on the trapped entity. There is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved. Here, we simulate the Brownian motion of a dimer and calculate the resulting scattered light, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals given a choice of scattering measurement angles.

@@

\section{Methodology}
\label{sec:Method}
\subsection{Brownian Simulation}
\label{sec:2.1}
We use the Brownian OT package developed by Jerome Fung \textit{et al} \cite{Vigilante2020Brownian_OT} to simulate the motion of a asymmetric dimer within an optical trap. Brownian OT is unique in that it combines MSTM \cite{Mishchenko1996MSTM} and 'Optical Tweezer Toolbox' (\textit{ott}) \cite{Lenton2020} to simulate the motion of arbitrary shaped sphere clusters. We simulated the motion of silica dimer trapped in a 5 mW Gaussian beam; the dimer was comprised of two spheres with radii $1\mu m$ and $ 0.5  \mu m$ respectively. We ran our simulation over the span of 10 seconds, with Brownian OT calculating the dimer's orientation and position every $1 \ ms$.

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{./Images/Dimer.png}
	\caption{Example dimer in orientation $\hat{s}$}
	\label{fig:dimer}
\end{figure}

With any trapped particle there are two main sources of motion, Brownian and optical forces. The dimer's Brownian motion ($\Delta q^{\textbf{B}}$) is simulated by generating a set of normally-distributed numbers whose covariance is given by the dimer's diffusion tensor $\textbf{D}$. The diffusion tensor is computed based on Farsund's work \cite{Farsund1996}. For computing the optical force Brownian OT uses the T-Matrix method to calculate the scattering coefficients of the outgoing scattered wave. The T-matrix describes the scattering process by relating the scattering wave to the incident wave:

@@

MSTM computes the T-matrix of the dimer, and then passes the T-matrix to \textit{ott} to compute the optical forces. The benefit of this approach is that MSTM can accurately compute the T-matrix for any arbitrary collection of spheres once, and then \textit{ott} can calculate the optical forces and torques. The total displacement is the sum of the optical and Brownian contributions:

@@

where \textbf{F} is the a 6 component vector that describes the optical force and torque components. Brownian OT returns the dimer's motion as a \textit{csv} file containing for each timestep the $[x,y,z]$ position of the dimer's centre of mass relative to the trap focus, and the orientation represented as a quaternion.

\subsection{Orientation estimation from scattering measurements}
\label{sec:2.2}

We represent the instantaneous orientation of the moving dimer in the trap as a three component unit vector,  $\hat{s}$, pointing from the centre of the larger sphere to the centre of the smaller sphere. A plane wave 'probe' laser beam directed perpendicular to the trapping laser, incident on the dimer, generates a scattering pattern dependent on the dimer's instantaneous orientation,  $I(\hat{s},\ \theta)$. This scattering pattern was computed using MSTM for each time step of our Brownian simulation. To represent a simulated 'measurement', three angles ($\theta_1,\ \theta_2,\ \& \theta_3$) were chosen and the calculated  intensity at each angle k recorded as $I(\hat{s},\ \theta_k)$.

Our goal is to estimate the dimer orientation $\hat{s}$ over time, based on the scattering measurements  $I(\hat{s},\ \theta_k)$: the question then is what set of angles ($\theta_1,\ \theta_2,\ \& \theta_3$) give us the most useful information on $\hat{s}$. To initially simplify the problem, rather than trying to get an exact estimation of $\hat{s}$ we discretise the orientation space into 30 possible reference orientations $\hat{n}_i$ that divide up a unit sphere evenly \cite{Rey2006}: our question becomes, how can we optimise the experiment to identify the reference orientation $\hat{n}_i$ closest to the real orientation $\hat{s}$. (Ultimately our method could be extended to an arbitrarily large number of reference orientations but for clarity and practicability we use a relatively small number here; additionally, the choice of a small number of discrete reference or 'candidate' orientations also accounts for the fact that any experiment will necessarily have a limited precision.)   Using MSTM we compute the scattering that would be produced by the dimer if it were in each of the reference orientations, denoted as $I(\hat{n}_i, \ \theta)$ and obtain 'measurements'  $I(\hat{n}_i, \ \theta_k)$ at our chosen set of  angles $\theta_k$. These raw intensities are normalised according to \eqref{eq:scale}:

\begin{align}
	\label{eq:scale}
	y_k(\hat{n}_i,\ \theta_k) = 
	 \frac{\left[I(\hat{n}_i, \ \theta_k)- \langle I(\hat{n}, \ \theta_k) \rangle \right]} 
	{\langle I^2(\hat{n},\ \theta_k) \rangle -\langle I(\hat{n}, \ \theta_k)\rangle^2}
\end{align}

Where the denominator is simply the standard deviation for our values of $I(\hat{n}, \theta_k)$. The dimer's \emph{actual}  instantaneous scattered intensity based on the real instantaneous orientation $\hat{s}$  is similarly scaled using \eqref{eq:scale} to give $y(\hat{s},\ \theta_k)$. See Table's A.1 and A.2 for numerical values of our reference orientation, raw intensities, and scaled signals. In practice $y(\hat{s}\, \theta_k)$ will have an inherent amount of uncertainty in our measurement due to noise from our detector. We compare our instantaneous signal to our reference signal by assuming that the measurement uncertainty for each detection fibre is Gaussian and is centred about $y_k(\hat{s})$ Combining all three detection angles provides us with the conditional probability that, should the dimer be in orientation $\hat{n}_i$, it would produce a signal $y(\hat{s})$:

\begin{align}
	\label{eq:conditional}
	p_i(y(\hat{s})\parallel\hat{n}_i) = \Pi^3_{k=1}
	(2\pi\sigma_k^2)^{-1/2} 
	e^{-(y_{k}(\hat{n}_i)-y_{k}(\hat{s}))^2/2\sigma_k^2}
\end{align}

Where $\sigma_k$ is the average error in the detected signal from our dimer $y_k(\hat{s})$. We left this as a tuneable variable to measure how the influence of experimental error would effect our predictive model. The signal error was simulated in our model by stochastically adjusting our values of $y_k(\hat{s})$.

@@

Where $\epsilon$ is the percent error expected from our detection fibres. Equation~\eqref{eq:conditional} assigns a probability to each reference orientation based on how similar the 'reference orientation signal' $y_{k}(\hat{n}_i)$ is to the actual signal $y_{k}(\hat{s})$ generated by the dimer in its real orientation $\hat{s}$. However, if our values of $y_k(\hat{n}_i)$ are similar to one another the probability distribution from equation~\ref{eq:conditional} will therefore assign similar probability for multiple reference orientations. Bayesian inference allows one to update a probability distribution by accounting for information known prior to making a measurement. We apply this by accounting for our distribution of reference orientations and instantaneous signals.

@@

Where $p_i(\hat{n})$ and $p(y)$ are our estimations of the prior distributions of particle orientations and instantaneous signals respectively. The former can be estimated by some Boltzmann distribution based on how far apart reference orientations are in physical space, covered in Section~\ref{sec:2.3}. The latter prior $p(y)$ is the probability of our dimer producing a signal $y_k(\hat{s})$ that falls with our distribution of reference signals $y_k(\hat{n})$. This is given by taking the discrete integral over collection of reference orientations.

@@

The implementation of \eqref{eq:bayes} improves our estimate by suppressing reference orientations that are far away from our dimer's orientation vector in physical space. In order to evaluate the accuracy of our model we tested it using our Brownian simulation of the dimer, see Section~\ref{sec:2.1}.

@@

\section{Results $\&$ Discussion}
\label{sec:Discussion}
\subsection{Asymmetric dimer dynamics}
\label{sec:3.1}
The Brownian OT software was used to simulate the motion of a trapped dimer ($a_1=1\mu m, a_2=0.5\mu m$) over the first 10 seconds of entering the optical trap. The initial orientation was assumed as strictly vertical (in line with the beam propagation direction). The dimer's position orientation was recorded every $10 \mu s$ for using as a test dataset for our model.

\begin{figure}[h]
	\label{fig:motion}
	\centering
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width =\textwidth]{./Images/traj.png}
	\end{subfigure}
	\begin{subfigure}{0.45\textwidth}
		\subcaption{}
		\includegraphics[width=\textwidth]{./Images/pos.png}
	\end{subfigure}
	\caption{Simulation results of: a) the dimers orientation vector with time, b) the dimer's [x,y,z] position with time.}
\end{figure}

@@

\subsection{Impact of signal error on model predictions}
\label{sec:3.2}
A key assumption of equation~\ref{eq:conditional} is that we assume our detected light signal will inherently have some degree of noise. This is the central premise that allows us to form equation~\eqref{eq:conditional}, understanding the limits of acceptable signal error is necessary for implementing this to experimental work.
To this end we ran our model while varying the value of $\epsilon$. \figurename{ 3} shows the model's performance with a selection of angles $15^{\circ},\ 55^{\circ}, \ \& \ 90^{\circ}$ and $\beta$ set at 1. The dotted line indicates the minimum radian distance ($0.137$ radians) between two neighbouring reference orientations, if we are under this line then we know our model is just off from selecting the correct orientation

\begin{figure}[h]

	\centering
	\begin{subfigure}{0.49\textwidth}
		\includegraphics[width=\textwidth]{./Images/epsilon_5.png}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
		\includegraphics[width=\textwidth]{./Images/epsilon_15.png}
	\end{subfigure}
	\begin{subfigure}{0.5\textwidth}
		\includegraphics[width=\textwidth]{./Images/epsilon_25.png}
	\end{subfigure}
	\caption{Model prediction for signal error of a) $5\%$, b) $15\%$, and c) $25\%$. Blue line denotes the best result we can achieve, orange line denotes the result provided by eq~\ref{eq:bayes}}
	\label{fig:epsilon}
\end{figure}

@@

\begin{figure}
	\label{fig:beta}
	\centering
	\begin{subfigure}{0.49\textwidth}
		\includegraphics[width=\textwidth]{./Images/beta_5.png}
	\end{subfigure}
	\begin{subfigure}{0.49\textwidth}
		\includegraphics[width=\textwidth]{./Images/beta_6.png}
	\end{subfigure}
	\caption{Model prediction for signal error of $25\%$ when a) $\beta = 5$ and b) $\beta=6$}
\end{figure}

As shown by \ref{fig:beta} eventually $\beta$ begins to over correct our estimate to the extent that we are a full radian away from the correct result. The exact value of $\beta$ that will optimise our performance varies for each choice of angle, therefore we can use it as variable for finely tuning our results to suit our choice of angles and our signal error. In the future we want to also investigate the possibility of extending our prior $p(\hat{n})$ so we account for not just the change in orientation after one time step but after 2 or 3 time steps to investigate if we can investigate the auto correlation of our dimer's motion.

\subsection{Parameter Optimisation Results}
\label{sec:3.3}
We wanted to see if our choice of angles plays a significant role in our model's efficacy. We employed ultranest to analyse the parameter space for our three angles by sampling across the priori of $10^{\circ} \ \& \ 180^{\circ}$. We kept both $\beta$ and $\sigma$ constant during this so we could as these are parameters that can be changed more freely.

\begin{figure}[h]
	\label{fig:corner}
	\centering
	\includegraphics[width=0.45\textwidth]{./Images/corneranglesfreed-1.png}
	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement}
\end{figure}

Figure~\ref{fig:corner} is the resulting corner plot from our ultranest sampler, darker areas indicate a choice of angles where the model performed better where as lighter areas are where the model did poorer. Overall from the corner plot provided it's clear that there is no direct correlation between our model's performance and the choice of detection angle. While the choice of angles does not provide a direct improvement in our estimation it does determine how finely we can estimate the orientation, for choices of angles that are close together the minimum spacing between our signals is so small that we can't achieve any greater accuracy than dividing the orientation space into 30 reference orientations. We hope to develop our model further in the future by being able to determine the maximum accuracy given that we know the signal error of our measurements.

@@

\section{Conclusion}
\label{sec:Conclusion}
We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern. In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity. The MSTM package is flexible in calculating the light scattering for a variety of micro-particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy. This has potential to be applied to characterise particle size and shape that are changing while in the optical trap. The authors would like to acknowledge and thank the support for this research from the funding provided by the Leverhulme Trust.\\
